---
title: "Weight Lifting Manner Prediction"
author: "Jackie Yang"
date: "Wednesday, October 15, 2014"
output: html_document
---

### Background

The data for this project come from accelerometers on the belt, forearm, arm, and dumbbell of 6 participants, who were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website [here](http://groupware.les.inf.puc-rio.br/har). We will build a machine learning algorithm to predict the manner in which they did the exercise based on a training data set. The model will be evaluated by predicting 20 different test cases. 

### Data Loading

```{r}
# Load libraries
library(caret)
library(corrplot)
# Download data
#if(!file.exists("data")) {
        #dir.create("data")
#}
#download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", "./data/train.csv")
#download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", "./data/test.csv")
#list.files("./data")
#date()
# Load Data
train <- read.csv("./data/train.csv", header = TRUE, na.strings = c("NA", ""))
test <- read.csv("./data/test.csv", header = TRUE, na.strings = c("NA", ""))
dim(train)
```

### Data Cleaning

```{r}
# Handle missing values
missing <- colSums(is.na(train))
unique(missing)
sum(missing == 19216)
train <- train[, missing == 0]
test <- test[, missing == 0]
dim(train)
# Remove unrelevant variables
names(train)
train <- train[, -c(1:7)]
test <- test[, -c(1:7)]
# Make sure the predictors are numeric
str(train$classe)
sum(sapply(train, is.numeric))
```

### Variable Selection

```{r}
# Remove highly correlated variables (r > 0.5)
corVar <- findCorrelation(cor(train[, -length(train)]), cutoff = 0.5)
train <- train[, -corVar]
test <- test[, -corVar]
corrplot(cor(train[, -length(train)]), method = "circle", type = "lower")
```

We have **`r ncol(train)-1`** predictors left in the training dataset, which is a good number for model fitting. 

### Data Partition and Model Fitting

```{r}
# Partition the training dataset in 70% training and 30% cross validation
inTrain <- createDataPartition(train$classe, p = 0.7, list = FALSE)
trainSub <- train[inTrain,]
crossVal <- train[-inTrain,]
# Fit the training data using Random Forests algorithm for accuracy
modFit <- train(classe ~ ., method = "rf", data = trainSub, 
                trControl = trainControl(method = "cv"), importance = TRUE)
# Plot predictor importance
varImpPlot(modFit$finalModel, sort = TRUE, type = 1, pch = 19, col = "blue")
```

### Model Validation

```{r}
pred <- predict(modFit, crossVal)
confMat <- confusionMatrix(pred, crossVal$classe)
confMat$table
```

This model has an accuracy of **`r confMat$overall[[1]]`** and an out-of-sample error of **`r 1-confMat$overall[[1]]`**.

### Test Data Prediction

```{r}
answers <- predict(modFit, test)
answers
# Create submission files
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
if(!file.exists("answers")) {
        dir.create("answers")
}
setwd("./answers")
pml_write_files(answers)
```

### Conclusion

We built a Random Forest machine learning model to predict weight lifting manners. This model has an accuracy of **`r confMat$overall[[1]]`** and an out-of-sample error of **`r 1-confMat$overall[[1]]`**. After submitting results to coursera, we got all the predictions right.